Improving legitimacy and efficiency \\ of decisions in democratic communities

### 2017-09-19 statistical-quorum.tex
Survey data.

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a novel concept of \emph{statistical quorum}
as a threshold of votes required to deem results legitimate.
\item We implemented an online application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases legitimacy of decisions,
improves utilization of collective attention,
and increases number of decisions made while not sacrificing their legitimacy.
\end{itemize}

@@

>>>>

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a novel concept of \emph{statistical quorum}
as a threshold of votes required to deem results legitimate.
\item We implemented an online application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases legitimacy of decisions,
improves utilization of collective attention, increases overall participation,
and increases number of decisions made while not sacrificing their legitimacy.
\end{itemize}

@@

To evaluate statistical quorum in practice, we developed an online web application PeerMind and
used it in a democratically run community of 140 members for 16 weeks, replacing the traditional
decision making through in-person meetings. The community is a house of 140
students living together and making collective decisions about the governance and operation of
the house, use of resources like common spaces and budgets, and changes to the house itself.

Before the introduction of the application, community had a practice of meeting weekly, with all
members being invited to participate. For decision making it used variation of Robert's rules~\cite{roberts}
with majority and super-majority voting to adopt decisions, generally using hand voting and voice voting.
A fixed quorum of the third of the membership was used. Meetings generally lasted between one to three hours.
Public minutes were taken of those meetings.

We introduced the application in fall 2016 semester for 16 weeks. We use spring 2016 as a baseline for comparison,
16 weeks as well, when traditional decision making was used.

To evaluate general perception of members before and after the use of the application for decision making,
we did a general survey at the end of each semester, asking members to answer general question about how
they felt the decision making in the house was that semester.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making in the house.
For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members
of the community were invited to participate in the survey.}
\label{fig:survey-comparison}
\end{figure}
